---
title: "Rgee: Demo"
output:
  html_document:
    df_print: paged
vignette: |
  %\VignetteIndexEntry{1_rgee_hello_world}  %\VignetteEngine{knitr::rmarkdown}  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  error = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

## 1) Load libraries
```{r, message=FALSE}
library(rgee)
library(sf)
library(cptcity)
library(mapview)
library(leaflet)
library(rnaturalearth)
library(reticulate)
```

Initialize earthengine
```{r}
ee_Initialize(drive = TRUE)
```

## 2) Study area

Defining study area with sf object
```{r}
world_map <- ne_countries(returnclass = "sf")
world_map <- world_map[world_map$name=='Ecuador',]
plot(world_map[1])
ee_Ecuador <- world_map %>% st_geometry() %>% sf_as_ee(check_ring_dir = TRUE)
```

## 3) Search into the Earth Engine Data Catalog

Defining study area with sf object
```{r}
ee_dataset() %>%
  ee_search_type("ImageCollection") %>%
  ee_search_tagstitle("landsat", "toa", "l8", logical_operator = "AND") %>%
  '['(1,'id') -> l8_name
```

## 4) Call the EE ImageCollection
```{r}
ic_l8 <- ee$ImageCollection(l8_name)$
  filterBounds(ee_Ecuador)$
  filterDate("2015-01-01", "2017-12-31")
```

## 5) Add NDVI band (FLOAT) into the ImageCollection
```{r}
addNDVI <- function(image) {
  ndvi = image$normalizedDifference(c('B5', 'B4'))$rename('NDVI');
  return(image$addBands(ndvi))
}
```

## 6) Quality mosaic vs Simple mean
```{r}
vizparams <- list(max=0.3, bands=c('B4', 'B3', 'B2'))

ic_l8_mean <- ic_l8$map(addNDVI)$mean()$clip(ee_Ecuador)
ic_l8_mosaic <- ic_l8$map(addNDVI)$qualityMosaic('NDVI')$clip(ee_Ecuador)

map_1 <- ee_map(ic_l8_mean,vizparams = vizparams,objname = 'mean')
map_2 <- ee_map(ic_l8_mosaic,vizparams = vizparams,objname = 'quality')
map_1 + map_2
```

## 7) Fast Download (< 5mb)
Download EE thumbnail images and read them as stars objects
```{r}
ecuador_stars <- ee_as_thumbnail(ic_l8_mosaic,vizparams = vizparams)
image(ecuador_stars, rgb = c(3,2,1))
```

## 6. Download using Google Chrome (> 5mb)
Resampling the image for a more faster download (this step could be omitted)
```{r}
image_to_download <- ic_l8_mosaic$
  select(c('B4','B3','B2'))$
  reproject(crs = "EPSG:4326", scale = 2500)
```

Passing from Earth Engine to Google Drive
```{r}
task_img <- ee$batch$Export$image$toDrive(
  image = image_to_download,
  folder = "ecuador_mosaic",
  fileFormat = "GEOTIFF",
  fileNamePrefix = "ecuador_mosaic"
)
```

ee_manage_cancel_all_running_taks()
```{r}
task_img$start()
ee_monitoring() # Monitoring task progress (Optional)
```

Passing from Google Drive to Hard disk
```{r}
ecuador_stars <- ee_download_drive(task_img)
plot(ecuador_stars)
image(ecuador_stars, rgb = c(1,2,3))
```