---
title: "Frequently Asked Questions"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here you can find possible **solutions for common `rgee` problems**. We will add to this page as soon as additional problems are reported. If these solutions do not work for you, or you have a different problem, signal it at: https://github.com/csaybar/rgee/issues

### Installation problem

#### ee_create_pyenv take by default a Python 2 version (ERROR: rgee just run under Python 3.5 >=)

It is a common scenario mostly for Windows users that have a lot of Python Envs setting on PATH. For solve this problem, take into account the follows steps:

```{r eval=FALSE}
  # 1. Find manually the Python Env that you want to use, for instance:
  myPythonPath <- "C:/Python36/ArcGISx6410.4/python.exe"
  
  # 2. Set the RETICULATE PYTHON PATH (restart R to see changes).
  ee_set_pyenv(myPythonPath, install = TRUE)

  # 3. Create a Python environment, e.g. ee.
  myPythonEnv <- ee_create_pyenv(python_env = "ee")

  # 4. Set the RETICULATE_PYTHON and RETICULATE_PYTHON_ENV (restart R to see changes).
  ee_set_pyenv(myPythonEnv, install = TRUE)
  
  # 5. Install Python package dependencies (restart R to see changes).
  ee_install_python_packages()
  
  # 6. Initialize rgee again!
  ee_Initialize()
```  
