---
title: "Frequently Asked Questions"
output: 
  github_document:
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here you can find possible **solutions for common `rgee` problems**. We will add to this page as soon as additional problems are reported. If these solutions does not work for you, or you have a different problem, signal it at: https://github.com/csaybar/rgee/issues

### Installation problem

### Python Configuration

- **`reticulate` do not recognize virtual environemnts.**

Some users could find works with virtual environments a struggle. The easiest solution to avoid this is to create a `.Renviron` file to store system variables. In this file we set the reticulate Python version we want to use and R first looks there when it is initiated. To create `.Renviron`, run as follows:

```{r Renviron}
reticulate_dir  <- path.expand("~/.Renviron")
fileConn <- file(reticulate_dir)
if (Sys.info()[['sysname']] == 'Linux') {
  writeLines('RETICULATE_PYTHON="/usr/bin/python3"', fileConn)
} else if (Sys.info()[['sysname']] == 'Windows') {
  writeLines('RETICULATE_PYTHON="C:\\Python35"', fileConn)  
} else if (Sys.info()[['sysname']] == 'Darwin') {
  writeLines('RETICULATE_PYTHON="/usr/bin/python3"', fileConn)
}
close(fileConn)
```

After that, it is just necessary to restart R.

### Selenium Configuration
