---
title: "World Seasonal NDVI - Animation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rgee_animation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes a simple pipeline to generate a world choropleth NDVI seasonal map. It tries to demostrate how easy is integrate `rgee` with the R spatial ecosystem.

### 1. Load R packages

```{r libraries , message=FALSE, warning=FALSE}
library(sf)
library(rgee)
library(stars)
library(dplyr)
library(lwgeom)
library(mapview)
library(cptcity)
library(ggplot2)
library(rnaturalearth)
library(gganimate)
```

### 2. Initialize Earth Engine

```{r init}
ee_Initialize()
```

### 3. Datasets

```{r datasets}
ndvi_palette <- cpt(pal = "cb_div_RdYlGn_03") # palette
modis <- ee$ImageCollection("MODIS/006/MCD43A4")
world_map <- ne_countries(returnclass = "sf")
land <- ee$FeatureCollection("USDOS/LSIB_SIMPLE/2017")
```

### 4. Functions

`calc_ndvi`: Calculate the NDVI from a MODIS image.

```{r calc_ndvi}
calc_ndvi <- function(img) {
  ndvi <- img$normalizedDifference(
    c("Nadir_Reflectance_Band2", 
      "Nadir_Reflectance_Band1")
  )
  select_NDVIband <- img$
    addBands(ndvi$rename("ndvi"))$
    select("ndvi")
  return(select_NDVIband)
}
```


`create_orto_basemap`: Create a world orthographic view in R. This function
is a simple adaptation of the original code of [Florian Zenoni](https://gist.github.com/fzenoni/ef23faf6d1ada5e4a91c9ef23b0ba2c1) (thanks a lot!).

```{r f2}
create_orto_basemap <- function(world_map, ortho_proj,lat) {
  circle <- st_point(x = c(0, 0)) %>%
    st_buffer(dist = 6371000) %>%
    st_sfc(crs = ortho_proj)
  circle_longlat <- circle %>% st_transform(crs = 4326)
  
  # circle_longlat cannot be used as it is
  # You must decompose it into a string with ordered longitudes
  # Then complete the polygon definition to cover the hemisphere
  if (lat != 0) {
    circle_longlat <- st_boundary(circle_longlat)
    
    circle_coords <- st_coordinates(circle_longlat)[, c(1, 2)]
    circle_coords <- circle_coords[order(circle_coords[, 1]), ]
    circle_coords <- circle_coords[!duplicated(circle_coords), ]
    
    circle_longlat <- st_linestring(circle_coords) %>% st_sfc(crs = 4326)
    
    # Rebuild line
    if (lat > 0) {
      rectangle <- list(rbind(
        circle_coords,
        c(X = 180, circle_coords[nrow(circle_coords), "Y"]),
        c(X = 180, Y = 90),
        c(X = -180, Y = 90),
        c(X = -180, circle_coords[1, "Y"]),
        circle_coords[1, c("X", "Y")]
      )) %>%
        st_polygon() %>%
        st_sfc(crs = 4326)
    } else {
      rectangle <- list(rbind(
        circle_coords,
        c(X = 180, circle_coords[nrow(circle_coords), "Y"]),
        c(X = 180, Y = -90),
        c(X = -180, Y = -90),
        c(X = -180, circle_coords[1, "Y"]),
        circle_coords[1, c("X", "Y")]
      )) %>%
        st_polygon() %>%
        st_sfc(crs = 4326)
    }
    circle_longlat <- st_union(st_make_valid(circle_longlat), st_make_valid(rectangle))
  }
  
  # A small negative buffer is necessary to avoid polygons still disappearing in a few pathological cases
  # I should not change the shapes too much
  visible <- st_intersection(world_map, st_buffer(circle_longlat, -0.09)) %>%
    st_transform(crs = ortho_proj)
  broken_reason <- st_is_valid(visible, reason = TRUE)
  
  # First fix NA's by decomposing them
  # Remove them from visible for now
  na_visible <- visible[is.na(broken_reason), ]
  visible <- visible[!is.na(broken_reason), ]
  
  # Open and close polygons
  na_visible <- st_cast(na_visible, "MULTILINESTRING") %>%
    st_cast("LINESTRING", do_split = TRUE)
  na_visible <- na_visible %>% mutate(npts = npts(geometry, by_feature = TRUE))
  # Exclude polygons with less than 4 points
  na_visible <- na_visible %>%
    filter(npts >= 4) %>%
    select(-npts) %>%
    st_cast("POLYGON")
  
  # Fix other broken polygons
  broken <- which(!st_is_valid(visible))
  for (land in broken) {
    result <- tryCatch({
      # visible[land,] <- st_buffer(visible[land,], 0) # Sometimes useful sometimes not
      visible[land, ] <- st_make_valid(visible[land, ]) %>%
        st_collection_extract()
    }, error = function(e) {
      visible[land, ] <<- st_buffer(visible[land, ], 0)
    })
  }
  
  # Bind together the two tables
  visible <- rbind(visible, na_visible)
  visible %>%
    "[["("geometry") %>%
    st_buffer(.00001) %>%
    st_union(quiet = TRUE) -> visible
  
  output_ortho <- list(world_limits = circle, layer = visible)
  return(output_ortho)
}
```

### 5. Global parameters

In this section, we will create the basemap (create_orto_basemap). You might change the global parameters LAT and LON according to personal preferences.

```{r Global parameters, message=FALSE, warning=FALSE}
LAT <- 10 
LON <- -80
ORTHO <- "+proj=ortho +lat_0=%s +lon_0=%s +x_0=0 +y_0=0 +a=6371000 +b=6371000 +units=m +no_defs"
ORTHO_PROJ <- sprintf(ORTHO, LAT, LON)
WORLD_MAP <- create_orto_basemap(world_map, ORTHO_PROJ, LAT)
```

### 6. Filtering and grouping the MODIS images to download

```{r modisdb}
modis_ndvi <- modis$filterDate("2010-01-01", "2016-01-01")$map(calc_ndvi)
djf <- modis_ndvi$filter(ee$Filter$calendarRange(12, 3, "month"))$mean()
mam <- modis_ndvi$filter(ee$Filter$calendarRange(3, 6, "month"))$mean()
jja <- modis_ndvi$filter(ee$Filter$calendarRange(6, 9, "month"))$mean()
son <- modis_ndvi$filter(ee$Filter$calendarRange(9, 12, "month"))$mean()

# Export as a stars object
visParams <- list(min = 0, max = 1, bands = "ndvi")
region <- c(-180, -90, 180, 90)
```


### 7. Exporting thumbnail images

The Earth Engine API offers the possibility to create thumbnail images. Inspired in the [cartoee Python Package](https://github.com/KMarkert/cartoee) (Developed by [Kel Markert](https://github.com/KMarkert/)), `rgee` implements [ee_as_thumbnail](https://csaybar.github.io/rgee/reference/ee_as_thumbnail.html). It is wrapper function around `ee$Image()$getThumbURL()`.

```{r message=FALSE,warning=FALSE}
modis_djf_stars <- ee_as_thumbnail(
  x = djf, 
  vizparams = visParams, 
  region = region) %>%
  '['(world_map) %>% 
  st_as_sf() %>% 
  st_transform(ORTHO_PROJ) %>% 
  st_make_valid() %>% 
  'colnames<-'(c("NDVI", "geometry"))

modis_mam_stars <- ee_as_thumbnail(
  x = mam, 
  vizparams = visParams, 
  region = region) %>%
  '['(world_map) %>% 
  st_as_sf() %>% 
  st_transform(ORTHO_PROJ) %>% 
  st_make_valid() %>% 
  'colnames<-'(c("NDVI", "geometry"))

modis_jja_stars <- ee_as_thumbnail(
  x = jja, 
  vizparams = visParams, 
  region = region) %>%
  '['(world_map) %>% 
  st_as_sf() %>% 
  st_transform(ORTHO_PROJ) %>% 
  st_make_valid() %>% 
  'colnames<-'(c("NDVI", "geometry"))

modis_son_stars <- ee_as_thumbnail(
  x = son, 
  vizparams = visParams, 
  region = region) %>%
  '['(world_map) %>% 
  st_as_sf() %>% 
  st_transform(ORTHO_PROJ) %>% 
  st_make_valid() %>% 
  'colnames<-'(c("NDVI", "geometry"))

modis_djf_stars$season <- "DJF"
modis_mam_stars$season <- "MAM"
modis_jja_stars$season <- "JJA"
modis_son_stars$season <- "SON"

modis_season <- rbind(
  modis_djf_stars, 
  modis_mam_stars,
  modis_jja_stars,
  modis_son_stars
)

```

### 7. Final step - gganimate

For create the world choropleth NDVI seasonal map, we put everything together. We will use `ggplot2` and `gganimate`:

```{r gganimate}
ggplot(modis_season) +
  geom_sf(
    data = WORLD_MAP$world_limits,
    fill = "aliceblue"
  ) +
  geom_sf(lwd = 0, aes(fill = NDVI)) +
  scale_fill_gradientn(colours = ndvi_palette) +
  labs(
    title = "NDVI - MCD43A4",
    subtitle = "Season:{current_frame}",
    caption = "Earth Engine Snippet: ee.ImageCollection('MODIS/006/MCD43A4')"
  ) +
  geom_sf(data = WORLD_MAP$layer, fill = NA) +
  coord_sf(crs = ORTHO_PROJ) +
  theme(
    panel.grid.major = element_line(
      color = gray(.7), linetype = "dashed", size = 0.5
    ),
    panel.ontop = TRUE,
    panel.background = element_rect(fill = NA)
  ) +
  # Here comes the gganimate code
  transition_manual(frames = season) +
  enter_fade() +
  exit_shrink() +
  ease_aes("linear")
```


