---
title: "Setup and Details"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


`rgee` has two types of dependencies. The first group called <span style="color:#b52b09">**strict dependencies**</span> that must be satisfied 
before the `rgee` installation. If this is not fulfilled 
**rgee just will not work**. The dependencies that comprised this group are:

- <span style="color:#b52b09"><b>
    Google account with Earth Engine activated
  </b></span>
- <span style="color:#b52b09"><b>
    Python > v3.5
  </b></span>
- <span style="color:#b52b09"><b>
    EarthEngine Python API
  </b></span>

The activation of **Earth Engine accounts** depends on each users, check
the oficial website of [Google Earth Engine](https://earthengine.google.com/) 
for more details. If you do not count with a **Python environment** in 
your system, run `ee_Initialize()` to display an interactive menu to install [Miniconda](https://docs.conda.io/en/latest/miniconda.html) (a free minimal 
installer for conda). We highly recommend use an virtual environment to use
`rgee`, you can create one using `ee_create_pyenv`. Finally, the
`ee_install_python_packages()` function is provided as a convenient way to
install all Python packages needed in `rgee`, even though it is not mandatory. 
You can count on with your own custom installation. This would be also allowed.
All the step described are summarised below:

```{r eval=FALSE}
library(rgee)
# Recommended way to install external rgee dependencies 

# 1. Initialize rgee with ee_Initialize(). If there is no any
#    Python environment, miniconda will be installed by default.
ee_Initialize()

# 2. Create a Python environment, e.g. ee.
pyenv <- ee_create_pyenv(python_env = "ee")

# Find others Python environments in the system.
# ee_discover_pyenvs()

# 3. Set a Python environment (e.g. ee) and restart R to see changes.
ee_set_pyenv(pyenv, install = TRUE)

# 4. Install Python package dependencies and restart R to see changes.
ee_install_python_packages()

# 5. Initialize rgee again!
ee_Initialize()
```


The second group of dependencies also called 
<span style="color:#857e04"><b>I/O dependencies</b></span> unlock `rgee` 
import & export functions. This dependencies are not mandatory. However, many
of the rgee I/O functionality depend on it. The dependencies that comprised this
group are shown below:

- <span style="color:#857e04">**Google Cloud Storage credentials**</span>
- <span style="color:#857e04">**Google Drive credentials**</span>

### Import and Export Spatial Data using rgee

The batch import/export involves difficulties for most GEE users. In `rgee`, we are aware of it and we created several functions to help users to download and upload spatial data. If you are trying to **download** data from GEE using `rgee` you will have three options:

- **ee_as_sf**: Useful to download ee$Geometry, ee$Feature and ee$FeatureCollection.
- **ee_as_***: Convert ee$Image to stars and raster object.
- **ee_as_thumbnail**: Create a raster or stars object based on a thumbnail image.

These functions through the argument **`via`** will permit you to **change the download method**. There are three methods to download an EE object in rgee: the `getInfo` which seeks the spatial object in the response, the `drive` options which use Google Drive (GD) as an container, and finally `gcs`  which use Google Cloud Storage (GCS). `rgee` to deal with **GD** and **GCS** use the R package [googledrive](https://googledrive.tidyverse.org/) and  [googleCloudStorageR](http://code.markedmondson.me/googleCloudStorageR/) respectively, so you will need to do the installation before.

```{r eval=FALSE}
# please try as follow
install.packages('googledrive')
install.packages('googleCloudStorageR')
```

**GD** is more friendly to novice Earth Engine users because the authentication process could be done without leaving R. However, if you are trying to move large amounts of data, it is preferable use Google Cloud Storage instead. It is important mentioning that, you will need to have your own Google Project with a credit card added to use GCS, charges will apply. See the [GCS_AUTH_FILE](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial to create your own service account key. If you want to understand why this is necessary, please have a look [Mark Edmondson](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial.

Batch **upload** is a harder process, in `rgee` we try to make it simple. If you want to upload files in a batch way, firstly you must **get authorization to read & write into a Google Cloud Storage (GCS) bucket**. `rgee` implement two functions to upload files:

- **sf_as_ee**: Convert a sfg, sfc or sf object to ee$FeatureCollection.
- **stars_as_ee**: Convert an stars object to ee$Image.
- **raster_as_ee**: Convert an raster object to ee$Image.

### Authorization and authentication

`rgee` deal with three different Google API's:

- Google Earth Engine
- Google Drive
- Google Cloud Storage

To authenticate either Google Drive or Google Cloud Storage, you just need to run as follow:

```{r eval=FALSE}
library(rgee)
#ee_reattach() # reattach ee as a reserve word

# Initialize just Earth Engine
ee_Initialize()
ee_Initialize(email = 'csaybar@gmail.com') # Use the argument email is not mandatory

# Initialize Earth Engine and GD
ee_Initialize(email = 'csaybar@gmail.com', drive = TRUE)

# Initialize Earth Engine and GCS
ee_Initialize(email = 'csaybar@gmail.com', gcs = TRUE)

# Initialize Earth Engine, GD and GCS
ee_Initialize(email = 'csaybar@gmail.com', drive = TRUE, gcs = TRUE)
```

If the Google account is verified and the permission is granted, you will be directed to an authentication token. Copy this token and paste it in the emerging GUI. This process will be repeated for each API, except for [Google Cloud Storage](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) see the link for details of how create a workflow that use GCS. If there is a successful outcome, all credentials are stored in:

```{r eval=FALSE}
ee_get_earthengine_path()
```

For the next sessions, there will not be need for any further authentication.

### Checking

The `ee_check()` function will help you for checking the sanity of `rgee` installation. Additionally, you can separately check the `rgee` dependencies with the following functions:

- `ee_check_python()` - Python version
- `ee_check_credentials()` - Google Drive and GCS credentials
- `ee_check_rgee_python_packages()` - Python packages

```{r eval=FALSE}
library(rgee)
ee_check()
```
