---
title: "Setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rgee_install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`rgee` has two types of dependencies. The <span style="color:#b52b09">**first group**</span> that must be satisfied before the `rgee` installation, and the <span style="color:#f0df26">**other one**</span> that is just necessary to unlock the `rgee` import & export functions. For instance, you will need firstly activate your Cloud Storage JSON API to use [ee_download_gcs](https://csaybar.github.io/rgee/reference/ee_download_gcs.html). All the external dependencies of `rgee` are shown bellow:

### External Required 

- <span style="color:#b52b09">Google acount with Earth Engine activated</span>
- <span style="color:#b52b09">Python > v3.5</span>
- <span style="color:#f0df26">Google Cloud Storage (GCS)</span>
- <span style="color:#f0df26">Google Chrome version > v76 </span>

### Third-Party Python Package Required

- <span style="color:#b52b09"> EarthEngine Python API (recommend version 0.1.175)</span>
- <span style="color:#f0df26"> selenium </span>
- <span style="color:#f0df26">beautifulsoup4</span>
- <span style="color:#f0df26">requests_toolbelt</span>

The `ee_install_rgee_python_packages()` function is provided as a convenient way to install all the necessary Python packages, but is not required. If you have an existing installation of the [Earth Engine Python API](https://pypi.org/project/earthengine-api/) or just prefer your own custom installation that’s fine too.

```{r eval=FALSE}
library(rgee)
# Recommended way to install Python packages
## 1. Create a virtualenv
virtualenv_create("rgee", python = "python3.7")
use_virtualenv("rgee")
ee_install_rgee_python_packages()
```

### Import and Export Spatial Data in GEE

The batch import/export involves a struggle for most GEE users. In `rgee` we are aware of that and created functions to help users to download and upload spatial data. If you are trying to **download** data from GEE, we highly recommend to use either [ee_download_drive](https://csaybar.github.io/rgee/reference/ee_download_drive.html) or [ee_download_gcs](https://csaybar.github.io/rgee/reference/ee_download_gcs.html). These functions depends on the [googledrive](https://googledrive.tidyverse.org/) and  [googleCloudStorageR](http://code.markedmondson.me/googleCloudStorageR/) R package respectively, so you will need installed it before.

```{r eval=FALSE}
# please try as follow
install.packages('googledrive')
install.packages('googleCloudStorageR')
```

Google Drive (`ee_download_gcs`) is more friendly with novice Earth Engine users because it just would need an authentication process (that is 100% automatic!). If you are trying to move large amounts of data, use Google Cloud Storage (`ee_download_gcs`) instead. However, you will need to have your own Google Project with a credit card added to used it, charges will apply. See the [GCS_AUTH_FILE](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial to create your own service account key. If you want to understand why this is necessary, please take a look to the [Mark Edmondson
](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial.

Batch **upload** is a harder process, in `rgee` we try to make simple through the [ee_upload](https://csaybar.github.io/rgee/reference/ee_upload.html) function. Always that you want to upload files in a batch way, will be necessary get authorization to read & write into a Google Cloud Storage (GCS) bucket. Fortunately, Earth Engine provides a **provisional** for free space into GCS through the URI gs://earthengine-uploads/ to all users. Inspired in the [geeup](https://pypi.org/project/geeup/0.3.3/) Python package we provide a connection to gs://earthengine-uploads/ when users do not defined the bucket argument in the `ee_upload` function. This process will use **Selenium** and require **ChromeDriver**. To install ChromeDriver in your system run as follow:

```{r eval=FALSE}
# Find the Google Chrome version of this system
# in chrome://settings/help
# In my case my Google Chrome version is 77.0.3865.75, so:
rgee::ee_install_drivers(GoogleChromeVersion = 77)
```

### Credentials

The `rgee` manage three different Google API's:

- Google Earth Engine
- Google Drive
- Google Cloud Storage

To authenticate the three APIs, you just need to run as follow:

```{r eval=FALSE}
library(rgee)
#ee_reattach() # reattach ee as a reserve word
ee_Initialize(drive = TRUE,gcs = TRUE)
```

If the Google account is verified and the permission is granted, you will be directed to an authentification token. Copy this token and paste it in the emerging GUI. This process will be repeated for each API, except for [Google Cloud Storage](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html). If the function runs successfully, all credentials are stored in:

```{r eval=FALSE}
path.expand("~/.config/earthengine/")
```

For the next sessions, there will be no need for any further authentification.

### Checking

The `ee_check()` function will help you for checking the sanity of your `rgee` installation. Additionally, you can separately check the rgee dependencies with the following functions:

- `ee_check_drivers()` - Chromedriver
- `ee_check_python()` - Python version
- `ee_check_credentials()` - Google Drive and GCS credentials
- `ee_check_rgee_python_packages()` - R/Python packages

```{r eval=TRUE}
library(rgee)
ee_check()
```
