---
title: "Setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rgee_install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`rgee` has two types of dependencies. The <span style="color:#b52b09">**first one**</span> must be satisfied before the `rgee` installation, and the <span style="color:#f0df26">**other one**</span> unlock `rgee` import & export functions. For instance, you will need firstly activate your Cloud Storage JSON API to use [ee_download_gcs](https://csaybar.github.io/rgee/reference/ee_download_gcs.html). All the external dependencies of `rgee` are shown below:

### External Required 

- <span style="color:#b52b09">Google acount with Earth Engine activated</span>
- <span style="color:#b52b09">Python > v3.5</span>
- <span style="color:#f0df26">Google Cloud Storage (GCS)</span>
- <span style="color:#f0df26">Google Chrome version > v76 </span>

### Third-Party Python Package Required

- <span style="color:#b52b09"> EarthEngine Python API</span>
- <span style="color:#f0df26"> selenium </span>
- <span style="color:#f0df26">beautifulsoup4</span>
- <span style="color:#f0df26">requests_toolbelt</span>

The `ee_install_rgee_python_packages()` function is provided as a convenient way to install all Python packages needed, even though it is not required.You can count on an existing installation of [Earth Engine Python API](https://pypi.org/project/earthengine-api/) or you can count on your own custom installation, all would be allowed.

```{r eval=FALSE}
library(rgee)
# Recommended way to install Python packages
## 1. Create a virtualenv
virtualenv_create("rgee", python = "python3.7")
use_virtualenv("rgee")
ee_install_rgee_python_packages()
```

### Import and Export Spatial Data in GEE

The batch import/export involves difficulties for most GEE users. In `rgee`, we are aware of it and we created functions to help users to download and upload spatial data. If you are trying to **download** data from GEE, we highly recommend to use either [ee_download_drive](https://csaybar.github.io/rgee/reference/ee_download_drive.html) or [ee_download_gcs](https://csaybar.github.io/rgee/reference/ee_download_gcs.html). These functions depends on [googledrive](https://googledrive.tidyverse.org/) and  [googleCloudStorageR](http://code.markedmondson.me/googleCloudStorageR/) R package respectively, so you will need to do the installation before.

```{r eval=FALSE}
# please try as follow
install.packages('googledrive')
install.packages('googleCloudStorageR')
```

Google Drive (`ee_download_drive`) is more friendly to novice Earth Engine users because the authentification process could be done without leaving R. However, if you are trying to move large amounts of data, it is preferable use Google Cloud Storage (`ee_download_gcs`) instead. It is important mentioning that, you will need to have your own Google Project with a credit card added to used it, charges will apply. See the [GCS_AUTH_FILE](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial to create your own service account key. If you want to understand why this is necessary, please have a look [Mark Edmondson](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html) tutorial.

Batch **upload** is a harder process, in `rgee` we try to make it simple by using the [ee_upload](https://csaybar.github.io/rgee/reference/ee_upload.html) function. Whether you want to upload files in a batch way, you must get authorization to read & write into a Google Cloud Storage (GCS) bucket. Fortunately, Earth Engine provides a **provisional** for free space to all users into GCS (gs://earthengine-uploads/). Inspired in the [geeup](https://pypi.org/project/geeup/0.3.3/) Python package we provide a connection to gs://earthengine-uploads/ when the bucket argument in the `ee_upload` function is not defined by the user. This process will use **Selenium** and will require **ChromeDriver**. To install ChromeDriver in your system run as follow:

```{r eval=FALSE}
# Find the Google Chrome version of this system
# in chrome://settings/help
# In my case my Google Chrome version is 77.0.3865.75, so:
rgee::ee_install_drivers(GoogleChromeVersion = 77)
```

### Credentials

`rgee` deal with three different Google API's:

- Google Earth Engine
- Google Drive
- Google Cloud Storage

To authenticate these three APIs, you just need to run as follow:

```{r eval=FALSE}
library(rgee)
#ee_reattach() # reattach ee as a reserve word
ee_Initialize(drive = TRUE,gcs = TRUE)
```

If the Google account is verified and the permission is granted, you will be directed to an authentication token. Copy this token and paste it in the emerging GUI. This process will be repeated for each API, except for [Google Cloud Storage](http://code.markedmondson.me/googleCloudStorageR/articles/googleCloudStorageR.html). If there is a successful outcome, all credentials are stored in:

```{r eval=FALSE}
path.expand("~/.config/earthengine/")
```

For the next sessions, there will not be need for any further authentication.

### Checking

The `ee_check()` function will help you for checking the sanity of `rgee` installation. Additionally, you can separately check the `rgee` dependencies with the following functions:

- `ee_check_drivers()` - Chromedriver
- `ee_check_python()` - Python version
- `ee_check_credentials()` - Google Drive and GCS credentials
- `ee_check_rgee_python_packages()` - R/Python packages

```{r eval=TRUE}
library(rgee)
ee_check()
```
